Enforced Invariant Confluence

### 2018-02-20 doc/paper_redux/new_statebased_sufficient.tex
Finished state based sufficient proof.

\begin{claim}
  Consider a state-based object $(S, s_0, \join)$, a set of transactions $T$,
  and an invariant $I$. $T$ is \Iconfluent{} if the following criteria are met:
  \begin{enumerate}
    \item
      $(S, \join)$ is a semilattice.
    \item
      For every $t \in T$, there exists some $t': S \to S$ such that for all
      $s \in S$, $t(s) = s \join t'(s)$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, $t_1(t_2(s)) = t_2(t_1(s))$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.
  \end{enumerate}
\end{claim}
\begin{proof}
  foo
\end{proof}

\todo{State conditions}
\todo{Prove sufficient}
\todo{Show more general than naive conditions}

>>>>

\begin{claim}
  Consider a state-based object $(S, s_0, \join)$, a set of transactions $T$,
  and an invariant $I$. $T$ is \Iconfluent{} if the following criteria are met:
  \begin{enumerate}
    \item
      $(S, \join)$ is a semilattice.

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that $t(s) = s
      \join s_t$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.
  \end{enumerate}
\end{claim}
\begin{proof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}

  To show that $T$ is \Iconfluent{}, it suffices to show that $T$ is
  \dIconfluent{} and \dIreducible{} (by
  \clmref{diconfluence_direducible_implies_iconfluence}). We begin by showing
  that $T$ is \dIconfluent{}. Let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. For ease
  of notation, let
  \[
    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
    \quad\text{and}\quad
    \baru{i} = u_j(\ldots(u_1(s_0))\ldots)
  \]
  We prove by strong induction on $m \in \nats$ that if $m = i + j$ where $0
  \leq i \leq n$ and $0 \leq j \leq m$, then $\Irec{\bart{i}(\baru{j}(s_0))}$.
  From this, it's immediate that $\Irec{\bart{n}(\baru{m}(s_0))}$, so
  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dIconfluent{}.
  \begin{itemize}
    \item \textbf{Case $m = 0$.}
      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ because
      $\Irec{e_1}$.

    \item \textbf{Case $m = 1$.}
      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
      \[
        \bart{1}(\baru{0}(s_0))
          = t_1(s_0)
      \]
      $\Irec{t_1(s_0)}$ because $\Irec{e_1}$.

    \item \textbf{Case $m \geq 2$.}
      If $i = 0$, then $j = m$ and $\bart{0}(s_0) \join \baru{j}(s_0) =
      \baru{j}(s_0)$ which recursively satisfies $I$ because $\Irec{e_2}$.
      Similarly if $j = 0$, then $i = m$ and $\bart{i}(s_0) \join \baru{0}(s_0)
      = \bart{i}(s_0)$ which recursively satisfies $I$ because $\Irec{e_1}$.
      %
      Otherwise, $i, j > 1$. Let
      \begin{align*}
        e_{i-1,j-1}
          &= \bart{i-1}(\baru{j-1}(s_0)) \\
        e_{i,j-1}
          &= \bart{i}(\baru{j-1}(s_0))
           = t_i(e_{i-1,j-1}) \\
        e_{i-1,j}
          &= \bart{i-1}(\baru{j}(s_0))
           = u_j(e_{i-1,j-1})
      \end{align*}
      By the inductive hypothesis, $\Irec{e_{i-1,j-1}}$, $\Irec{e_{i,j-1}}$,
      and $\Irec{e_{i-1,j}}$. By condition 3, $e_{i-1,j} \join e_{i,j-1} =
      \bart{i}(s_0) \join \baru{j}(s_0) = \bart{i}(\baru{j}(s_0))$ satisfies
      $I$. This implies that $\Irec{\bart{i}(\baru{j}(s_0))}$.
  \end{itemize}

  Next, we show that $T$ is \dIreducible{}. Again let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free expressions that recursively satisfy $I$. We just
  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
  expression, so $T$ is \dIreducible{}.
\end{proof}
